Benchmarking Python Deep Learning Frameworks\\ for Language Modeling on GPUs

### 2016-12-14 paper/main.tex
Cleanup, add final dynet results, draft conclusion

\begin{abstract}
Neural networks are omnipresent in natural language processing. 
We benchmark three popular Python frameworks (DyNet, TensorFlow and Theano) on a realistic NLP task: language modeling.
\ms{Change if needed:} We find that DyNet is significantly faster on this task. However, performance wasn't the only bottleneck during development, and there are other variables to consider.
\end{abstract}

@@

Traditionally, neural models are trained until their performance on a \textit{development} set stops improving. Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
\section{Benchmarks}
\label{sec:benchmarks}
\subsection{Time}
The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie DyNet is the best (so far) (2 sentences perhaps)}.
\begin{table}
\begin{tabular}{c|ccc}
Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
Test epoch 		& 3m4s			& 7m45s 			& 24m \\
Train batch 		& 1s 				& 5s 					& 1s \\
Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
\end{tabular}
\caption{\label{tab:timing}Total runtime on various subtasks.}
\end{table}

@@

As discussed in section \ref{sec:benchmarks}, we found that DyNet was the fastest at performing our specific language modeling task. \lhl{TODO: flesh this out with a sentence or two more -- maybe something like ``DyNet seems to make certain optimizations that are advantageous in this setting''}
DyNet seems to make certain optimizations that are advantageous in this setting, perhaps specifically in the memory management choices it makes.\ms{is this okay?}

@@

\subsubsection{Ease of use}
\lhl{TODO: all the things that TensorFlow/DyNet support out of the box.
comment somewhere about documentation, debuggability\ms{Done?}}

\paragraph{DyNet}
DyNet includes built-in constructors for various NLP-relevant models, including several varieties recurrent neural nets like Tree LSTMs and encoder-decoders. This makes it easy for the programmer to build many common kinds of models. DyNet also allows for dynamic computation graph, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths. While our neural language model used a static computation graph, DyNet made graph construction easy.

@@

\section{Conclusion}
\ms{Many factors, no clear winner; }

>>>>

\begin{abstract}
Neural networks are omnipresent in natural language processing. 
We benchmark three popular Python frameworks (DyNet, TensorFlow and Theano) on a realistic NLP task: language modeling.
We find that DyNet is significantly faster on this task. However, performance wasn't the only bottleneck during development, and there are other variables to consider.
\end{abstract}

@@

Traditionally, neural models are trained until their performance on a \textit{development} set stops improving. Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
\section{Benchmarks}
\label{sec:benchmarks}
\subsection{Time}
The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. Overall, DyNet seems to converge the fastest of the three; its epoch time is dramatically shorter, but it also took more epochs to converge.

\begin{table}
\begin{tabular}{c|ccc}
Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
Convergence		& 26h21m 		& 35h25m 			& (TBD)\\
Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
Test epoch 		& 3m4s			& 7m45s 			& 24m \\
Train batch 		& 1s 				& 5s 					& 1s \\
Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
\end{tabular}
\caption{\label{tab:timing}Total runtime on various subtasks.}
\end{table}

@@

As discussed in section \ref{sec:benchmarks}, we found that DyNet was the fastest at performing our specific language modeling task.
DyNet seems to make certain optimizations that are advantageous in this setting, perhaps specifically in the memory management choices it makes, or in adapting to inputs of different sizes.

@@

\subsubsection{Ease of use}

\paragraph{DyNet}
DyNet includes built-in constructors for various NLP-relevant models, including several varieties of recurrent neural nets like Tree LSTMs and encoder-decoders. This makes it easy for the programmer to build many common kinds of models. DyNet also allows for dynamic computation graph, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths. While our neural language model used a static computation graph, DyNet made graph construction easy.

@@

\section{Conclusion}
\ms{Many factors, no clear winner; }
We implemented a simple language modeling task in three deep learning frameworks, and measured their performance when running on a GPU. We found that DyNet converged faster than TensorFlow or Theano, but analyzing performance on a fine-grained level showed that all frameworks used the underlying GPU resources with comparable efficiency. Thus, efficiency is only one of several factors researchers should consider in choosing a deep learning framework. Users should also consider ease of installation and use, level of support available, and the suitability of the framework for their particular task.
